\documentclass[10pt, twocolumn]{article}
\usepackage{latex8}
\usepackage{times}
\usepackage{amsmath}
\usepackage{alltt}

\newtheorem{definition}{Definition}
\newtheorem{examp}{Example}
\newenvironment{example}{\begin{examp}\rm}{\rule{2mm}{2mm}\end{examp}}

\begin{document}
  \title{
    PolicyUpdater -- A System for Dynamic Access Control:  \\
    Formalization, Implementation and a Case Study
  }

  \author{
    Vino Fernando Crescini and Yan Zhang                   \\
    School of Computing and Information Technology         \\
    University of Western Sydney                           \\
    Penrith South DC, NSW 1797, Australia                  \\
    E-mail: \{jcrescin,yan\}@cit.uws.edu.au
  }

  \date{}

  \maketitle

  \begin{abstract}
    [text]
  \end{abstract}

  \section{Introduction}

    \subsection{Overview of Logic-Based Approaches}

    \subsection{Contributions}

    \subsection{Structure of This Paper}

  \section{Language $\cal{L}$}

    Language $\cal{L}$ is a first-order logic language that represents a policy
    base for an authorization system. Two key features of the language are: (1)
    providing a means to conditionally and dynamically update the existing
    policy base and (2) having a mechanism by which queries may be evaluated
    from the updated policy base.

    \subsection{Syntax}

      Logic programs of language ${\cal L}$ are composed of language
      statements, each terminated by a semicolon ";" character. C-style
      comments delimited by the "/*" and "*/" characters may appear anywhere in
      the logic program.

      \subsubsection{Components of Language $\cal{L}$}

        \paragraph{Identifiers.}
          The most basic unit of language $\cal{L}$ is the identifier.
          Identifiers are used to represent the different components of the
          language, and are divided into three main classes:

          \begin{itemize}
            \item
              {\em Entity Identifiers} represent constant entities that make up
              a logical atom. They are divided further into three types, with
              each type again divided into the {\em singular entity} and
              {\em group entity} categories:

              \begin{itemize}
                \item
                  {\em Subjects}: e.g. alice, lecturers, group.
                \item
                  {\em Access Rights}: e.g. read, write, own.
                \item
                  {\em Objects}: e.g. file, database, directory.
              \end{itemize}

              An entity identifier is defined as a lower-case alphabet
              character, followed by 0 to 127 characters of alphabet, digit or
              underscore characters. The following regular expression shows the
              syntax of entity identifiers:

              \begin{verbatim}[a-z]([a-zA-Z0-9_]{0,127}\end{verbatim}

            \item
              {\em Policy Update Identifiers} are used for the sole purpose of
              naming a policy update. These identifier names are then used as
              labels to refer to policy update definitions and directives. As
              labels, identifiers of this class occupy a different namespace
              from entity identifiers. For this reason, policy update
              identifiers share the same syntax with entity identifiers:

              \begin{verbatim}[a-z]([a-zA-Z0-9_]{0,127}\end{verbatim}
 
            \item
              {\em Variable Identifiers} are used as entity identifier
              place-holers, and as such, are subject to the entity identifier
              types. Variable identifiers are prefixed with 1 or 2 upper-case
              characters to distinguish them from entity identifiers, followed
              by a number composed of 1 to 6 digits. The upper-case characters
              determines the type of the variable identifier. The following is
              a list of regular expressions for each variable identifier type:

              \begin{itemize}
                \item
                  Subject variables (singular and group)

                  \begin{verbatim}S([0-9]){1,6}\end{verbatim}
                \item
                  Subject variables (singular)

                  \begin{verbatim}SS([0-9]){1,6}\end{verbatim}
                \item
                  Subject variables (group)

                  \begin{verbatim}SG([0-9]){1,6}\end{verbatim}
                \item
                  Access Right variables (singular and group)

                  \begin{verbatim}A([0-9]){1,6}\end{verbatim}
                \item
                  Access Right variables (singular)

                  \begin{verbatim}AS([0-9]){1,6}\end{verbatim}
                \item
                  Access Right variables (group)

                  \begin{verbatim}AG([0-9]){1,6}\end{verbatim}
                \item
                  Object variables (singular and group)

                  \begin{verbatim}O([0-9]){1,6}\end{verbatim}
                \item
                  Object variables (singluar)

                  \begin{verbatim}OS([0-9]){1,6}\end{verbatim}
                \item
                  Object variables (group)

                  \begin{verbatim}OG([0-9]){1,6}\end{verbatim}
              \end{itemize}
          \end{itemize}

        \paragraph{Atoms.}
          An atom is composed of a relation with 2 to 3 entity or variable
          identifiers that represent a logical relationship between the
          entities. There are three types of atoms:

          \begin{itemize}
            \item
              {\em Holds.} An atom of this type states that the subject
              identifier $sub$ holds the access right identifier $acc$
              for the object identifier $obj$.
         
              \begin{verbatim}holds(<sub>, <acc>, <obj>)\end{verbatim}
            \item
              {\em Membership.} This type of atom states that the singular
              identifier $elt$ is a member or element of the group identifier
              $grp$. It is important to note that identifiers $elt$ and $grp$
              must be of the same base type (e.g. subject and subject group).
         
              \begin{verbatim}memb(<elt>, <grp>)\end{verbatim}
            \item
              {\em $Subset$.} The subset atom states that the group identifiers
              $grp1$ and $grp2$ are of the same types and that group $grp1$ is
              a subset of the group $grp2$.

              \begin{verbatim}subst(<grp1>, <grp2>)\end{verbatim}
          \end{itemize}

          Atoms that contain no variables, i.e. composed entirely of entity
          identifiers, are called {\em ground atoms}.

        \paragraph{Facts.}
          A fact makes a claim that the relationship represented by an atom or
          its negation holds in the current context. Facts are negated by the
          use of the negation operator ($!$). The following shows the formal
          syntax of a fact:
 
          \begin{verbatim}[!]<holds_atm>|<memb_atm>|<subst_atm>\end{verbatim}

          Note that facts may be made up of atoms that contain variable
          identifiers. Facts with no variable occurences are called
          {\em ground facts}.
 
        \paragraph{Expressions.}
          An expression is either a fact, or a logical conjunction of facts,
          separated by the double-ampersand characters $\&\&$.

          \begin{verbatim}<fact1> [&& <fact2> [&& ...]]\end{verbatim}

          Expressions that are made up of only ground facts are called
          {\em ground expressions}.

      \subsubsection{Definition Statements}

        \paragraph{Entity Identifier Definition.}

          All entity identifiers (subjects, access rights, objects and groups)
          must first be declared before any other statements to define the
          entity domain of the policy base. The following entity declaration
          syntax illustrates how to define one or more entity identifiers of a
          particular type.

          \begin{verbatim}ident sub|acc|obj[-grp] <e_id>[, ...]];\end{verbatim}

        \paragraph{Initial Fact Definition.}

          The initial facts of the policy base, those that hold before any
          policy updates are performed, are defined by using the following
          definition syntax:

          \begin{verbatim}initially <ground-exp>;\end{verbatim}

        \paragraph{Constraint Definition.}

          Constraints are logical rules that holds regardless of any changes
          that may occur when the policy base is updated. The constraint rules
          are true in the initial state and remains true after any policy
          update.

          The constraint syntax below shows that for any state of the policy
          base, expression $exp1$ holds if expression $exp2$ is true and there
          is no evidence that $exp3$ is true. The $with$ $absence$ clause
          allows constraints to behave like default propositions, where the
          absence of proof that an expression holds implies that the default
          expression holds.

          It is important to note that the expressions $exp1$, $exp2$ and
          $exp3$ are non-ground expressions, which means identifiers within
          them may be variables.

          \begin{verbatim}
always <exp1>
  [implied by <exp2>
  [with absence <exp3>]];
          \end{verbatim}

        \paragraph{Policy Update Definition.}

          Before a policy update can be applied, it must first be defined by
          using the following syntax:

          \begin{verbatim}
<up_id>([<var_id>[, ...]])
  causes <exp1>
  if <exp2>;
          \end{verbatim}

          $up\_id$ is the policy update identifier to be used in referencing
          this policy update. The optional $var\_id$ list are the variable
          identifiers occurring the expressions $exp1$ and $exp2$ and will
          eventually be replaced by entity identifiers when the update is
          referenced. The postcondition expression $exp1$ is an expression that
          will hold in the state after this update is applied. The expression
          $exp2$ is a precondition expression that must hold in the current
          state before this update is applied.

          It is important to note that a policy update definition will have no
          effect on the policy base until it is appled by one of the directives
          described in the following section.

        \subsubsection{Directive Statements}

        \paragraph{Policy Update Directives.}

        The policy update sequence list contains a list of references to
        defined policy updates in the domain. The policy updates in the
        sequence list are applied to the current state of the policy base one
        at a time to produce a policy base state upon which queries can be
        evaluated. 

        The following four directives are the policy sequence manipulation
        features of language $\cal{L}$.

        \subparagraph{Adding an update into the sequence.}
          Defined policy updates are added into the sequence list through the
          use of the following directive:

          \begin{verbatim}seq add <up_id>([<e_id>[, ...]]);\end{verbatim}

          \noindent where $up\_id$ is the identifier of a defined policy
          update and the $e\_id$ list is a comma-separated list of entity
          identifiers that will replace the variable identifiers that occur in
          the definition of the policy update. 

        \subparagraph{Listing the updates in the sequence.}
          The following directive may be used to list the current contents of
          the policy update sequence list.

          \begin{verbatim}seq list;\end{verbatim}

          This directive is answered with an ordinal list of policy updates in
          the form:

          \begin{verbatim}<n> <up_id>([e_id[, ...]])\end{verbatim}

          \noindent where $n$ is the ordinal index of the policy update within
          the sequence list starting at 0. $up\_id$ is the policy update
          identifier and the $e\_id$ list is the list of entity identifiers
          used to replace the variable identifier place-holders.

        \subparagraph{Removing an update from the sequence.}
          The syntax below shows the directive to remove a policy update
          reference from the list. $n$ is the ordinal index of the policy
          update to be removed. Note that removing a policy update reference
          from the sequence list may change the ordinal index of other update
          references.

          \begin{verbatim}seq del <n>;\end{verbatim}

        \subparagraph{Computing an update sequence.}

          The policy updates in the sequence list does not get applied until
          the $compute$ directive is issued. The directive causes the policy
          update references in the sequence list to be applied one at a time in
          the same order that they appear in the list. The directive also
          causes the system to generate the policy base models against which
          query requests can be evaluated.

          \begin{verbatim}compute;\end{verbatim}

        \paragraph{Query Directive.}

          A ground query expression may be issued against the current state of
          the policy base. This current state is derived after all the updates
          in the update sequence have been applied, one at a time, upon the
          initial state. Query expressions are answered with a $true$, $false$
          or an $unknown$, depending on whether the queried expression holds,
          its negation holds, or neither, respectively. Syntax is as follows:

          \begin{verbatim}query <ground-exp>;\end{verbatim} 

        \begin{example}
          The following language ${\cal L}$ program code listing shows a simple
          rule-based document access control system scenario.

          In this example, the subject $alice$ initially is a member of the
          subject group $sudents$, which is holds a $read$ access right for the
          object $file$. The constraint states that if the $grp$ group
          has $read$ access for $file$, and no other information is present to
          conclude that $grp$ do not have $write$ access for $file$, then
          the $grp$ group is granted $write$ access for $file$.

          \begin{verbatim}
/* entity identifier definitions */

ident sub alice;
ident sub-grp grp;
ident acc read, write;
ident obj file;

/* initial fact definitions */

initially 
  memb(alice, grp) &&
  holds(grp, read, file);

/* constraint definitions */

always holds(grp, write, file)
  implied by
    holds(grp, read, file)
  with absence
    !holds(grp, write, file);

/* policy update definitions */

delete_read(SG0, OS0)
  causes !holds(SG0, read,  OS0);

/* sequence manipluation directives */

seq add delete_read(grp, file);
compute;

/* query directives */

query holds(grp, write, file);
query holds(grp, read, file);
query holds(alice, write, file);
query holds(alice, read, file);
          \end{verbatim}
        \end{example}

    \subsection{Semantics}

      \subsubsection{Domain Description of Language ${\cal L}$}

        \begin{definition}
          The domain description ${\cal D}_{\cal L}$ of language ${\cal L}$ is
          defined as a finite set of ground initial state facts, constraint
          rules and policy update definitions.
        \end{definition}

        In addition to the domain description ${\cal D}_{\cal L}$, language
        ${\cal L}$ also includes an additional ordered set: the sequence list
        $\psi$. The sequence list $\psi$ is an ordered set that contains a
        sequence of references to policy update definitions. Each policy update
        reference consists of the policy update identifier and a series of zero
        or more identifier entities to replace the variable place-holders in
        the policy update definitions.

      \subsubsection{Language ${\cal L}^{*}$}

        In language ${\cal L}$, the policy base is subject to change, which is
        triggered by the application of policy updates. Such changes bring
        forth the concept of policy base states. Conceptually, a state may be
        thought of as a set of facts and constraints of the policy base at a
        particular instant. The state transition notation below shows that a
        new state $PB'$ is generated from the current state $PB$ after the
        policy update $u$ is applied.
  
        \begin{quote}
          $PB$ $\overrightarrow{_{u}}$ $PB'$
        \end{quote}

        This definition of a state means that for every policy update applied
        to the policy base, a new instance of the policy base or a new set of
        facts and constraints are generated. To precisely define the underlying
        semantics of domain description ${\cal D}_{\cal L}$ in language
        ${\cal L}$, we introduce language ${\cal L}^{*}$, which is an extended
        logic program representation of langauge ${\cal L}$, with state as an
        explicit sort.

        Language ${\cal L}^{*}$ contains only one special state constant
        $S_{0}$ to represent the initial state of a given domain description.
        All other states are represented as a resulting state obtained by
        applying the $Result$ function.

        The $Result$ function takes a policy update reference $u$, where $u$
        $\in$ $\psi$, and the current state $\sigma$ as input arguments and
        returns the resulting state $\sigma'$ after update $u$ has been applied
        to state $\sigma$.
 
        \begin{quote}
          $\sigma'$ = $Result$($u$, $\sigma$)
        \end{quote}

        Given an initial state $S_{0}$ and a sequence list $\psi$, each state
        may be represented as:

        \begin{quote}
          $\sigma_{0}$ = $S_{0}$

          $\sigma_{1}$ = $Result$($u_{0}$, $\sigma_{0}$)

          ...

          $\sigma_{|\psi|}$ = $Result$($u_{|\psi| - 1}$, $\sigma_{|\psi| - 1}$)
        \end{quote}

        Substituting each state with a recursive call to the $Result$ function,
        the final state $S_{|\psi|}$ is defined as follows:

        \begin{quote}
          $S_{|\psi|}$ = $Result$($u_{|\psi| - 1}$, $Result$(..., $Result$($u_{0}$, $S_{0}$)))
        \end{quote}

        \paragraph{Entities.}

          The entity set ${\cal E}$ is the union of six disjoint entity sets:
          single subject ${\cal E}_{ss}$, group subject ${\cal E}_{sg}$,
          single access right ${\cal E}_{as}$, group access right
          ${\cal E}_{ag}$, single object ${\cal E}_{os}$ and group object
          ${\cal E}_{og}$. Each entity in set ${\cal E}$ corresponds directly
          to the \emph{entity identifiers} of language ${\cal L}$.

          \begin{enumerate}
            \item
              ${\cal E} = {\cal E}_{s} \cup {\cal E}_{a} \cup {\cal E}_{o}$
            \item
              ${\cal E}_{s} = {\cal E}_{ss} \cup {\cal E}_{sg}$
            \item
              ${\cal E}_{a} = {\cal E}_{as} \cup {\cal E}_{ag}$
            \item
              ${\cal E}_{o} = {\cal E}_{os} \cup {\cal E}_{og}$
          \end{enumerate}

        \paragraph{Atoms.}

          The main difference between language ${\cal L}$ and language
          ${\cal L}^{*}$ lies in the definition of an atom. Atoms in language
          ${\cal L}^{*}$ represent a logical replationship of two to three
          entities, as with atoms of language ${\cal L}$. Furthermore, atoms of
          language ${\cal L}^{*}$ extends this definition by defining the
          state of the policy base in which the relationship holds. In this
          paper, atoms of language ${\cal L}^{*}$ are written with the
          hat character ($\hat{holds}$, $\hat{memb}$ and $\hat{subst}$) to
          differentiate from the atoms of language ${\cal L}$.

          The atom set ${\cal A}^{\sigma}$ is the set of all atoms in state
          $\sigma$. 

          \begin{enumerate}
            \item
              ${\cal A}^{\sigma} = {\cal A}^{\sigma}_{h} \cup {\cal A}^{\sigma}_{m} \cup {\cal A}^{\sigma}_{s}$
            \item
              ${\cal A}^{\sigma}_{h} = \{\forall (s \in {\cal E}_{s}, a \in {\cal E}_{a}, o \in {\cal E}_{o}), \hat{holds}(s, a, o, \sigma) \}$
            \item
              ${\cal A}^{\sigma}_{m} = {\cal A}^{\sigma}_{ms} \cup {\cal A}^{\sigma}_{ma} \cup {\cal A}^{\sigma}_{mo}$
            \item
              ${\cal A}^{\sigma}_{s} = {\cal A}^{\sigma}_{ss} \cup {\cal A}^{\sigma}_{sa} \cup {\cal A}^{\sigma}_{so}$
            \item
              ${\cal A}^{\sigma}_{ms} = \{\forall (e \in {\cal E}_{ss}, g \in {\cal E}_{sg}), \hat{memb}(e, g, \sigma)\}$
            \item
              ${\cal A}^{\sigma}_{ma} = \{\forall (e \in {\cal E}_{as}, g \in {\cal E}_{ag}), \hat{memb}(e, g, \sigma)\}$
            \item
              ${\cal A}^{\sigma}_{mo} = \{\forall (e \in {\cal E}_{os}, g \in {\cal E}_{og}), \hat{memb}(e, g, \sigma)\}$
            \item
              ${\cal A}^{\sigma}_{ss} = \{\forall (g1, g2 \in {\cal E}_{sg}), \hat{subst}(g1, g2, \sigma)\}$
            \item
              ${\cal A}^{\sigma}_{sa} = \{\forall (g1, g2 \in {\cal E}_{ag}), \hat{subst}(g1, g2, \sigma)\}$
            \item
              ${\cal A}^{\sigma}_{so} = \{\forall (g1, g2 \in {\cal E}_{og}), \hat{subst}(g1, g2, \sigma)\}$
          \end{enumerate}

        \paragraph{Facts.}
       
          A fact is a logical statement that makes a claim that an atom either
          holds or does not hold at a particular state. The following is the
          formal definition of fact $f$ in state $\sigma$:

          \begin{quote}
            $f^{\sigma}$ = $[\lnot]$$\alpha$, $\alpha$ $\in$ ${\cal A}^{\sigma}$
          \end{quote}

        \paragraph{Expressions and Propositions.}
          Expressions or conjunctions of facts are represented as follows:

          \begin{quote}
            $f_{0}$, ..., $f_{n}$

            where $n$ $\geq$ $0$
          \end{quote}

          Propositions are implications composed of a single fact $c$ as the
          consequent and an expression $p$ as the premise.

          \begin{quote}
            $c$ $\leftarrow$ $p_{0}$, ..., $p_{n}$

            where $n$ $\geq$ $0$
          \end{quote}

      \subsubsection{Translating Language ${\cal L}$ to Language ${\cal L^{*}}$}

        \begin{definition}
          Given a domain description ${\cal D}_{\cal L}$ of language
          ${\cal L}$, $Tran$(${\cal D}_{\cal L}$) is defined as the
          extended logic program ${\cal L}^{*}$ translation of language
          ${\cal L}$. The domain description ${\cal D}_{\cal L}$ is said to be
          {\em consistent} if and only if $Trans$(${\cal D}_{\cal L}$) has
          a consistent answer set. 
        \end{definition}

        Before we can fully define the $Trans$(${\cal D}_{\cal L}$), we must
        first define the following functions:

        \paragraph{}

          Another function, $CopyAtom$, takes two arguments: an atom
          $\hat{\alpha}$ of language ${\cal L}^{*}$ at some state $\sigma$ and
          another state $\sigma'$. The function returns an equivalent atom of
          the same type and with the same entities, but in the new state
          specified.

          \begin{quote}
            $CopyAtom$($\hat{\alpha}$, $\sigma'$)
 
            = 
            \begin{math}
              \begin{cases}
                \mbox{$\hat{holds}$($s$, $a$, $o$, $\sigma'$), if $\hat{\alpha}$ = $\hat{holds}$($s$, $a$, $o$, $\sigma$)} \\
                \mbox{$\hat{memb}$($e$, $g$, $\sigma'$), if $\hat{\alpha}$ = $\hat{memb}$($e$, $g$, $\sigma$)} \\
                \mbox{$\hat{subst}$($g_{1}$, $g_{2}$, $\sigma'$), if $\hat{\alpha}$ = $\hat{subst}$($g_{1}$, $g_{2}$, $\sigma$)}
              \end{cases}
            \end{math}
          \end{quote}

        \paragraph{}

          The $TransAtom$ function takes an atom $\alpha$ of language
          ${\cal L}$ and an arbitrary state $\sigma$ and returns the equivalent
          atom of language ${\cal L}^{*}$.
 
          \begin{quote}
            $TransAtom$($\alpha$, $\sigma$)
 
            = 
            \begin{math}
              \begin{cases}
                \mbox{$\hat{holds}$($s$, $a$, $o$, $\sigma$), if $\alpha$ = $holds$($s$, $a$, $o$)} \\
                \mbox{$\hat{memb}$($e$, $g$, $\sigma$), if $\alpha$ = $memb$($e$, $g$)} \\
                \mbox{$\hat{subst}$($g_{1}$, $g_{2}$, $\sigma$), if $\alpha$ = $subst$($g_{1}$, $g_{2}$)}
              \end{cases}
            \end{math}
          \end{quote}

        \paragraph{}

          The $TransFact$ function is similar to the $TransAtom$ function, but
          instead of translating an atom, it takes a fact from language
          ${\cal L}$ and a state then returns the equivalent fact in language
          ${\cal L}^{*}$.

        \paragraph{Initial Fact Expression.}

          Translating initial fact expressions of language ${\cal L}$ to
          language ${\cal L}^{*}$ is a trivial procedure: translate each atom
          that make up the initial fact expression of language ${\cal L}$
          with its corresponding equivalent initial state atom of language
          ${\cal L}^{*}$. 

          Given an initial fact expression of language ${\cal L}$ that is
          composed of $n$ facts:

          \begin{alltt}
  initially \(f\sb{0}\) && ... && \(f\sb{n}\)
          \end{alltt}

          In language ${\cal L}^{*}$, the statement above is represented as:

          \begin{quote}
            $\hat{f_{0}}$ $\leftarrow$

            ...

            $\hat{f_{n}}$ $\leftarrow$

            where $\hat{f_{i}}$ $=$ $TransFact$($f_{i}$, $S_{0}$),
            $0$ $\leq$ $i$ $\leq$ $n$
          \end{quote}

        \paragraph{Constraint Rules.}

          Each constraint rule in language ${\cal L}$ is expressed as a series
          of logical rules in language ${\cal L}_{*}$. Given that all variable
          occurences have been grounded to entity identifiers, a constraint in
          language ${\cal L}$, with $m$, $n$, $o$ $\geq$ $0$ may be represented
          as:

          \begin{alltt}
  always \(a\sb{0}\) && ... && \(a\sb{m}\)
  implied by \(b\sb{0}\) && ... && \(b\sb{n}\)
  with absence \(c\sb{0}\) && ... && \(c\sb{o}\)
          \end{alltt}
         
          Each fact in the $always$ clause of language ${\cal L}$ corresponds
          to a new rule, where it is the consequent. Each of these new rules
          will have expression $b$ in the $implied$ $by$ clause as the positive
          premise and the expression $c$ in the $with$ $absence$ clause as the
          negative premise.

          \begin{quote}
            $a_{0}$ $\leftarrow$
            $b_{0}$, ..., $b_{n}$,
            $not$ $c_{0}$, ..., $not$ $c_{o}$

            ...

            $a_{m}$ $\leftarrow$
            $b_{0}$, ..., $b_{n}$,
            $not$ $c_{0}$, ..., $not$ $c_{o}$
          \end{quote}

          Under the definition of constraints, each of the rules above must be
          made to hold in all states as defined by the sequence list $\psi$.
          This can be accomplished by translating each of the rules above to
          a set of $|\psi|$ rules, one for each state.

           \begin{quote}
            $\hat{a}^{S_{0}}_{0}$ $\leftarrow$
            $\hat{b}^{S_{0}}_{0}$, ..., $\hat{b}^{S_{0}}_{n}$,
            $not$ $\hat{c}^{S_{0}}_{0}$, ..., $not$ $\hat{c}^{S_{0}}_{o}$

            ...

            $\hat{a}^{S_{|\psi|}}_{0}$ $\leftarrow$
            $\hat{b}^{S_{|\psi|}}_{0}$, ..., $\hat{b}^{S_{|\psi|}}_{n}$,
            $not$ $\hat{c}^{S_{|\psi|}}_{0}$, ..., $not$ $\hat{c}^{S_{|\psi|}}_{o}$

            ...

            $\hat{a}^{S_{0}}_{m}$ $\leftarrow$
            $\hat{b}^{S_{0}}_{0}$, ..., $\hat{b}^{S_{0}}_{n}$,
            $not$ $\hat{c}^{S_{0}}_{0}$, ..., $not$ $\hat{c}^{S_{0}}_{o}$

            ...

            $\hat{a}^{S_{|\psi|}}_{m}$ $\leftarrow$
            $\hat{b}^{S_{|\psi|}}_{0}$, ..., $\hat{b}^{S_{|\psi|}}_{n}$,
            $not$ $\hat{c}^{S_{|\psi|}}_{0}$, ..., $not$ $\hat{c}^{S_{|\psi|}}_{o}$

            where

            $\hat{a}^{\sigma}_{i}$ $=$ $TransFact$($a_{i}$, $\sigma$),
            $0$ $\leq$ $i$ $\leq$ $m$,

            $\hat{b}^{\sigma}_{j}$ $=$ $TransFact$($b_{j}$, $\sigma$),
            $0$ $\leq$ $j$ $\leq$ $n$,

            $\hat{c}^{\sigma}_{k}$ $=$ $TransFact$($c_{k}$, $\sigma$),
            $0$ $\leq$ $k$ $\leq$ $o$,

            $S_{0}$ $\leq$ $\sigma$ $\leq$ $S_{|\psi|}$
          \end{quote}

        \paragraph{Policy Updates.}

          With all occurences of variable placeholders grounded to entity
          identifiers, a policy update $u$, as defined by language ${\cal L}$
          is in the form:

          \begin{quote}
            $u$ causes $a_{0}$ \&\& ... \&\& $a_{m}$
            if $b_{0}$ \&\& ... \&\& $b_{n}$
          \end{quote}

          In language ${\cal L}^{*}$, such policy updates may be represented as
          a set of implications, with each fact $a$ in the postcondition
          expression as the consequent and precondition expression $b$ as the
          premise. However, the translation process must also take into account
          that the premise of the implication holds in the state before the
          policy update is applied and that the consequent holds in the state
          after the application.

          \begin{quote}
            $\hat{a}_{0}$ $\leftarrow$ $\hat{b}_{0}$, ..., $\hat{b}_{n}$

            ...

            $\hat{a}_{m}$ $\leftarrow$ $\hat{b}_{0}$, ..., $\hat{b}_{n}$

            where

            $\hat{a}_{i}$ $=$ $TransFact$($a_{i}$, $Result$($u$, $\sigma$)),
            $0$ $\leq$ $i$ $\leq$ $m$,

            $\hat{b}_{j}$ $=$ $TransFact$($b_{j}$, $\sigma$),
            $0$ $\leq$ $j$ $\leq$ $n$
          \end{quote}

        \paragraph{Additional Constraints.}

          In addition to the above translations, there are a few other implicit
          constraint rules implied by language ${\cal L}$ that needs to be
          explicitly defined in language ${\cal L}^{*}$.

          \subparagraph{Inheritance Rules.}

            All properties held by a group is inherited by all the members and
            subsets of that group. This rule is easy to apply for subject group
            entities. However, careful attention must be given to access right
            and object groups. A subject holding an access right for an object
            group implies that that subject also holds that access right for
            all objects in the object group. Similarly, a subject holding an
            access right group for a particular object implies that the subject
            holds all access rights contained in the access right group for
            that object.

            In a situation where a particular property is to be inherited by
            an entity from a group of which it is a member or subset, and the
            contained entity already holds the negation of that property, a
            conflict is encountered. This conflict is resolved by giving
            negative facts higher precedence over its positive counterpart:
            by allowing member or subset entities to inherit its parent group's
            properties only if the entities do not already hold the negation of
            those properties.

            The following are the inheritance constriant rules to allow the
            properties held by a group to propagate to its members and
            subsets that do not already hold the negation of the properties.

            \begin{enumerate}
              \item
                Subject Group Membership Inheritance

                \begin{math}
                  \begin{aligned}[t]
                    \hat{holds}(s_{s}, a, o, \sigma) \leftarrow &
                    \hat{holds}(s_{g}, a, o, \sigma), \\
                    & \hat{memb}(s_{s}, s_{g}, \sigma), \\
                    & not \lnot \hat{holds}(s_{s}, a, o, \sigma) \\
                    \lnot \hat{holds}(s_{s}, a, o, \sigma) \leftarrow &
                    \lnot \hat{holds}(s_{g}, a, o, \sigma), \\
                    & \hat{memb}(s_{s}, s_{g}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($s_{s}$, $s_{g}$, $a$, $o$, $\sigma$), where

                $s_{s} \in {\cal E}_{ss}$,
                $s_{g} \in {\cal E}_{sg}$,
                $a \in {\cal E}_{a}$,
                $o \in {\cal E}_{o}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$

              \item
                Access Right Group Membership Inheritance

                \begin{math}
                  \begin{aligned}[t]
                    \hat{holds}(s, a_{s}, o, \sigma) \leftarrow &
                    \hat{holds}(s, a_{g}, o, \sigma), \\
                    & \hat{memb}(a_{s}, a_{g}, \sigma), \\
                    & not \lnot \hat{holds}(s, a_{s}, o, \sigma) \\
                    \lnot \hat{holds}(s, a_{s}, o, \sigma) \leftarrow &
                    \lnot \hat{holds}(s, a_{g}, o, \sigma), \\
                    & \hat{memb}(a_{s}, a_{g}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($s$, $a_{s}$, $a_{g}$, $o$, $\sigma$), where

                $s \in {\cal E}_{s}$,
                $a_{s} \in {\cal E}_{as}$,
                $a_{g} \in {\cal E}_{ag}$,
                $o \in {\cal E}_{o}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$

              \item
                Object Group Membership Inheritance

                \begin{math}
                  \begin{aligned}[t]
                    \hat{holds}(s, a, o_{s}, \sigma) \leftarrow &
                    \hat{holds}(s, a, o_{g}, \sigma), \\
                    & \hat{memb}(o_{s}, o_{g}, \sigma), \\
                    & not \lnot \hat{holds}(s, a, o_{s}, \sigma) \\
                    \lnot \hat{holds}(s, a, o_{s}, \sigma) \leftarrow &
                    \lnot \hat{holds}(s, a, o_{g}, \sigma), \\
                    & \hat{memb}(o_{s}, o_{g}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($s$, $a$, $o_{s}$, $o_{g}$, $\sigma$) where

                $s \in {\cal E}_{s}$,
                $a \in {\cal E}_{a}$,
                $o_{s} \in {\cal E}_{os}$,
                $o_{g} \in {\cal E}_{og}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$

              \item
                Subject Group Subset Inheritance

                \begin{math}
                  \begin{aligned}[t]
                    \hat{holds}(s_{g1}, a, o, \sigma) \leftarrow &
                    \hat{holds}(s_{g2}, a, o, \sigma), \\
                    & \hat{subst}(s_{g1}, s_{g2}, \sigma), \\
                    & not \lnot \hat{holds}(s_{g1}, a, o, \sigma) \\
                    \lnot \hat{holds}(s_{g1}, a, o, \sigma) \leftarrow &
                    \lnot \hat{holds}(s_{g2}, a, o, \sigma), \\
                    & \hat{subst}(s_{g1}, s_{g2}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($s_{g1}$, $s_{g2}$, $a$, $o$, $\sigma$), where

                $s_{g1}, s_{g2} \in {\cal E}_{sg}$,
                $a \in {\cal E}_{a}$,
                $o \in {\cal E}_{o}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$

              \item
                Access Right Group Subset Inheritance

                \begin{math}
                  \begin{aligned}[t]
                    \hat{holds}(s, a_{g1}, o, \sigma) \leftarrow &
                    \hat{holds}(s, a_{g2}, o, \sigma), \\
                    & \hat{subst}(a_{g1}, a_{g2}, \sigma), \\
                    & not \lnot \hat{holds}(s, a_{g1}, o, \sigma) \\
                    \lnot \hat{holds}(s, a_{g1}, o, \sigma) \leftarrow &
                    \lnot \hat{holds}(s, a_{g2}, o, \sigma), \\
                    & \hat{subst}(a_{g1}, a_{g2}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($s$, $a_{g1}$, $a_{g2}$, $o$, $\sigma$), where

                $s \in {\cal E}_{s}$,
                $a_{g1}, a_{g2} \in {\cal E}_{ag}$,
                $o \in {\cal E}_{o}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$

              \item
                Object Group Subset Inheritance

                \begin{math}
                  \begin{aligned}[t]
                    \hat{holds}(s, a, o_{g1}, \sigma) \leftarrow &
                    \hat{holds}(s, a, o_{g2}, \sigma), \\
                    & \hat{subst}(o_{g1}, o_{g2}, \sigma), \\
                    & not \lnot \hat{holds}(s, a, o_{g1}, \sigma) \\
                    \lnot \hat{holds}(s, a, o_{g1}, \sigma) \leftarrow &
                    \lnot \hat{holds}(s, a, o_{g2}, \sigma), \\
                    & \hat{subst}(o_{g1}, o_{g2}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($s$, $a$, $o_{g1}$, $o_{g2}$, $\sigma$), where

                $s \in {\cal E}_{s}$,
                $a \in {\cal E}_{a}$,
                $o_{g1}, o_{g2} \in {\cal E}_{og}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$
            \end{enumerate}

          \subparagraph{Transitivity Rules.}

            Given three groups $G$, $G'$ and $G''$. If $G$ is a subset of $G'$
            and $G'$ is a subset of $G''$, then $G$ must also be a subset of
            $G''$. The following rules ensure that the transitive property of
            subject, access right and object groups hold:

            \begin{enumerate}
              \item
                Subject Group Transitivity

                \begin{math}
                  \begin{aligned}[t]
                    \hat{subst}(sg_{1}, sg_{3}, \sigma) \leftarrow &
                    \hat{subst}(sg_{1}, sg_{2}, \sigma), \\
                    & \hat{subst}(sg_{2}, sg_{3}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($sg_{1}$, $sg_{2}$, $sg_{3}$, $\sigma$), where

                $sg_{1}, sg_{2}, sg_{3} \in {\cal E}_{sg}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$

              \item
                Access Right Group Transitivity

                \begin{math}
                  \begin{aligned}[t]
                    \hat{subst}(ag_{1}, ag_{3}, \sigma) \leftarrow &
                    \hat{subst}(ag_{1}, ag_{2}, \sigma), \\
                    & \hat{subst}(ag_{2}, ag_{3}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($ag_{1}$, $ag_{2}$, $ag_{3}$, $\sigma$), where

                $ag_{1}, ag_{2}, ag_{3} \in {\cal E}_{ag}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$

              \item
                Object Group Transitivity

                \begin{math}
                  \begin{aligned}[t]
                    \hat{subst}(og_{1}, og_{3}, \sigma) \leftarrow &
                    \hat{subst}(og_{1}, og_{2}, \sigma), \\
                    & \hat{subst}(og_{2}, og_{3}, \sigma)
                  \end{aligned}
                \end{math}

                $\forall$ ($og_{1}$, $og_{2}$, $og_{3}$, $\sigma$), where

                $og_{1}, og_{2}, og_{3} \in {\cal E}_{og}$,
                $S_{0} \leq \sigma \leq S_{|\psi|}$
            \end{enumerate}

          \subparagraph{Inertial Rules.}

            Intuitively, all facts in the current state that are not affected
            by a policy update should be carried over to the next state after
            the update. In language ${\cal L}^{*}$, this rule must be
            explicitly stated as a constraint. Formally, the intertial rules
            are expressed as follows:

            \begin{quote}
              $\hat{\alpha}'$ $\leftarrow$ $\hat{\alpha}$, $not$ $\lnot$ $\hat{\alpha}'$

              $\lnot$ $\hat{\alpha}'$ $\leftarrow$ $\lnot$ $\hat{\alpha}$, $not$ $\hat{\alpha}'$

              $\forall$($\hat{\alpha}$,$u$), where 
              $\hat{\alpha}$ $\in$ ${\cal A}^{\sigma}$,
              $u$ $\in$ $\psi$,

              $\hat{\alpha}'$ = $CopyAtom$($\hat{\alpha}$, $Result$($u$, $\sigma$))
            \end{quote}

        \begin{example}
          The following shows the language ${\cal L}$ program listing in
          Example1,  translated into language ${\cal L}^{*}$.


          \begin{enumerate}
            \item
              Initial State Facts

              \begin{math}
                \hat{memb}(alice, grp, S_{0}) \\
                \hat{holds}(grp, read, file,S_{0})
              \end{math}

            \item
              Constraints

              \begin{math}
                  \begin{aligned}[t]
                    % constraints (S0)
                    \hat{holds}&(grp, write, file, S_{0}) \leftarrow \\
                    & \hat{holds}(grp, read, file, S_{0}), \\
                    & not \lnot \hat{holds}(grp, write, file, S_{0}) \\
                    % constraints (S1)
                    \hat{holds}&(grp, write, file, S_{1}) \leftarrow \\
                    & \hat{holds}(grp, read, file, S_{1}), \\
                    & not \lnot \hat{holds}(grp, write, file, S_{1})
                  \end{aligned}
              \end{math}

            \item
              Policy Update

              \begin{math}
                \lnot \hat{holds}(grp, read, file, S_{1}) \leftarrow
              \end{math}

            \item
              Inheritance Rules

              \begin{math}
                \begin{aligned}[t]
                  % inheritance rules (positive, read, S0)
                  \hat{holds}&(alice, read, file, S_{0}) \leftarrow \\
                  & \hat{holds}(grp, read, file, S_{0}), \\
                  & \hat{memb}(alice, grp, S_{0}), \\
                  & not \lnot \hat{holds}(alice, read, file, S_{0}) \\
                  % inheritance rules (negative, read, S0)
                  \lnot \hat{holds}&(alice, read, file, S_{0}) \leftarrow \\
                  & \lnot \hat{holds}(grp, read, file, S_{0}), \\
                  & \hat{memb}(alice, grp, S_{0})
                \end{aligned}
              \end{math}

              \begin{math}
                \begin{aligned}[t]
                  % inheritance rules (positive, write, S0)
                  \hat{holds}&(alice, write, file, S_{0}) \leftarrow \\
                  & \hat{holds}(grp, write, file, S_{0}), \\
                  & \hat{memb}(alice, grp, S_{0}), \\
                  & not \lnot \hat{holds}(alice, write, file, S_{0}) \\
                  % inheritance rules (negative, write, S0)
                  \lnot \hat{holds}&(alice, write, file, S_{0}) \leftarrow \\
                  & \lnot \hat{holds}(grp, write, file, S_{0}), \\
                  & \hat{memb}(alice, grp, S_{0})
                \end{aligned}
              \end{math}

              \begin{math}
                \begin{aligned}[t]
                  % inheritance rules (positive, read, S1)
                  \hat{holds}&(alice, read, file, S_{1}) \leftarrow \\
                  & \hat{holds}(grp, read, file, S_{1}), \\
                  & \hat{memb}(alice, grp, S_{1}), \\
                  & not \lnot \hat{holds}(alice, read, file, S_{1}) \\
                  % inheritance rules (negative, read, S1)
                  \lnot \hat{holds}&(alice, read, file, S_{1}) \leftarrow \\
                  & \lnot \hat{holds}(grp, read, file, S_{1}), \\
                  & \hat{memb}(alice, grp, S_{1})
                 \end{aligned}
              \end{math}

              \begin{math}
                \begin{aligned}[t]
                  % inheritance rules (positive, write, S1)
                  \hat{holds}&(alice, write, file, S_{1}) \leftarrow \\
                  & \hat{holds}(grp, write, file, S_{1}), \\
                  & \hat{memb}(alice, grp, S_{1}), \\
                  & not \lnot \hat{holds}(alice, write, file, S_{1}) \\
                  % inheritance rules (negative, write, S1)
                  \lnot \hat{holds}&(alice, write, file, S_{1}) \leftarrow \\
                  & \lnot \hat{holds}(grp, write, file, S_{1}), \\
                  & \hat{memb}(alice, grp, S_{1})
                \end{aligned}
              \end{math}

            \item
              Transitivity Rules
              % no transitivity rules

            \item
              Inertial Rules

              \begin{math}
                \begin{aligned}[t]
                  % inertial rule holds(alice, read, file)
                  \hat{holds}&(alice, read, file, S_{1}) \leftarrow \\
                  & \hat{holds}(alice, read, file, S_{0}), \\
                  & not \lnot \hat{holds}(alice, read, file, S_{1}) \\
                  % inertial rule !holds(alice, read, file)
                  \lnot \hat{holds}&(alice, read, file, S_{1}) \leftarrow \\
                  & \lnot \hat{holds}(alice, read, file, S_{0}), \\
                  & not \lnot \hat{holds}(alice, read, file, S_{1})
                \end{aligned}
              \end{math}

              \begin{math}
                \begin{aligned}[t]
                  % inertial rule holds(alice, write, file)
                  \hat{holds}&(alice, write, file, S_{1}) \leftarrow \\
                  & \hat{holds}(alice, write, file, S_{0}), \\
                  & not \lnot \hat{holds}(alice, write, file, S_{1}) \\
                  % inertial rule !holds(alice, write, file)
                  \lnot \hat{holds}&(alice, write, file, S_{1}) \leftarrow \\
                  & \lnot \hat{holds}(alice, write, file, S_{0}), \\
                  & not \lnot \hat{holds}(alice, write, file, S_{1})
                \end{aligned}
              \end{math}

              \begin{math}
                \begin{aligned}[t]
                  % inertial rule holds(grp, read, file)
                  \hat{holds}&(grp, read, file, S_{1}) \leftarrow \\
                  & \hat{holds}(grp, read, file, S_{0}), \\
                  & not \lnot \hat{holds}(grp, read, file, S_{1}) \\
                  % inertial rule !holds(grp, read, file)
                  \lnot \hat{holds}&(grp, read, file, S_{1}) \leftarrow \\
                  & \lnot \hat{holds}(grp, read, file, S_{0}), \\
                  & not \lnot \hat{holds}(grp, read, file, S_{1})
                \end{aligned}
              \end{math}

              \begin{math}
                \begin{aligned}[t]
                  % inertial rule holds(grp, write, file)
                  \hat{holds}&(grp, write, file, S_{1}) \leftarrow \\
                  & \hat{holds}(grp, write, file, S_{0}), \\
                  & not \lnot \hat{holds}(grp, write, file, S_{1}) \\
                  % inertial rule !holds(grp, write, file)
                  \lnot \hat{holds}&(grp, write, file, S_{1}) \leftarrow \\
                  & \lnot \hat{holds}(grp, write, file, S_{0}), \\
                  & not \lnot \hat{holds}(grp, write, file, S_{1})
                \end{aligned}
              \end{math}

              \begin{math}
                \begin{aligned}[t]
                  % inertial rule memb(alice, grp) 
                  \hat{memb}&(alice, grp, S_{1}) \leftarrow \\
                  & \hat{memb}(alice, grp, S_{0}), \\
                  & not \lnot \hat{memb}(alice, grp, S_{1}) \\
                  % inertial rule !memb(alice, grp) 
                  \lnot \hat{memb}&(alice, grp, S_{1}) \leftarrow \\
                  & \lnot \hat{memb}(alice, grp, S_{0}), \\
                  & not \hat{memb}(alice, grp, S_{1})
                \end{aligned}
              \end{math}
          \end{enumerate}
        \end{example}
      \subsection{Evaluation}

        \begin{definition}
          Given a consistent domain description ${\cal D}_{\cal L}$, ground
          query expression $\phi$ and a finite sequence list $\psi$, we say the
          query $\phi$ holds in ${\cal D}_{\cal L}$ after the policy updates
          in the sequence list $\phi$ has been applied:

          \begin{quote}
            ${\cal D}_{\cal L}$ $\models$ \{$\phi$, $\psi$\}
          \end{quote}

          \noindent if and only if for every atom $\alpha$ in $\phi$, $\alpha$
          $\in$ ${\cal A}^{|\psi|}$, [$\lnot$] $\alpha$ is in every answer
          set of $trans$(${\cal D}_{\cal L}$)
        \end{definition}

    \begin{example}
      [a full example (translation + result)]

    \end{example}

  \section{Semantic Properties}

  \section{Implementation}

    Although the semantics of language ${\cal L}$ shows that any program $P$
    of language ${\cal L}$ can be translated into an extended logic program
    $P^{*}$ of language ${\cal L}^{*}$, the stable model semantics requires 
    a normal logic program for computation. Therefore, a program $P$ of
    language ${\cal L}$ must then be further translated into a normal logic
    program first before any query evaluation or policy update can be
    performed.

    An extended logic program $P^{*}$ in language ${\cal L}^{*}$ is translated
    into a normal logic program by performing 2 additional steps:

    \begin{itemize}
      \item
        Removal of classical negation
      \item
        Representing facts in propositional form
    \end{itemize}

    The removal of classical negation from facts of language ${\cal L}$ is a
    straightforward procedure: each classically negated fact $\lnot$ $f$ is
    replaced by a new and unique positive fact $f'$ that represents the
    negation of fact $f$. Then, the following constraint rule must be added to 
    preserve the consistency of the policy base for all facts $f$ in the
    domain.

    \begin{quote}
      $FALSE$ $\leftarrow$ $f$, $f'$
    \end{quote}

    Semantically, an atom of language ${\cal L}^{*}$ is defined as a relational
    binding of 2 to 3 entity identifiers and a state parameter. To translate
    into a normal logic program, atoms of this first-order logic form must be
    converted into propositional form, that is, an atom must be represented by
    a single identity or symbol instead of a relation. Such translation is
    achieved by the following functions:

    \begin{itemize}
      \item
        $f^{NLP}$ = $Encode$($f^{\cal L}$)
      \item
        $f^{\cal L}$ = $Decode$($f^{NLP}$)
    \end{itemize}

    Our proposed translation method from language ${\cal L}$ to a normal logic
    program relies mainly on the representation of atoms and facts in the
    implementation. As we shall illustrate in the following sections, the
    representation of atoms and facts plays a major role in both the removal of
    classical negation and the translation to propositional form. However,
    before showing the translation, the data structures must first be defined.

    \subsection{Data Structures}

      As a program of language ${\cal L}$ is parsed, each statement containing
      entity declarations, facts, constraint rules and policy updates must
      be stored in a buffer structure before the translation process is
      started. This buffer structure is composed of the symbol table and the
      policy base.

      \subsubsection{Symbol Table}

        The symbol table is used to store the identifier entities defined in the
        entity identifier declaration section of language ${\cal L}$ programs.
        The symbol table is composed of 6 separate string lists:

        \begin{quote}
          \begin{tabular}[t]{|l|l|l|}
            \hline
            \textbf{Field} & \textbf{Type} & \textbf{Description} \\
            \hline
            $ss$ & string list & single subject \\
            \hline
            $sg$ & string list & group subject \\
            \hline
            $as$ & string list & single access \\
            \hline
            $ag$ & string list & group access right \\
            \hline
            $os$ & string list & single object \\
            \hline
            $og$ & string list & group object \\
            \hline
          \end{tabular}
        \end{quote}
 
        In addition to the above lists, the symbol table also defines 3
        pseudo-lists that are concatenations of the above lists:
 
        \begin{quote}
          \begin{tabular}[t]{|l|l|l|}
            \hline
            \textbf{Field} & \textbf{Type} & \textbf{Description} \\
            \hline
            $s$ & string list & $ss$ $+$ $sg$ \\
            \hline
            $a$ & string list & $as$ $+$ $ag$ \\
            \hline
            $o$ & string list & $os$ $+$ $og$ \\
            \hline
          \end{tabular}
        \end{quote}

        Each entity identifier are sorted in the above lists accoring to their
        type, and ordered accoring to the order in which they are declared in
        the program. Each list is indexed by consecutive positive integers
        starting from zero. Pseudo-lists are ordered in the same manner as the
        other lists, with the single list elements starting at index 0, and
        the group list elements offset by the length of the single list.

      \subsubsection{Policy Base}

        When a language ${\cal L}$ program is parsed, each of the facts, rules
        and policy updates must be stored into the policy base before they can
        be translated into language $\cal L^{*}$. The policy base is composed
        of 4 tables to store the following: initial state facts, constraint
        rules, policy update definitions and the policy update sequence.

        \paragraph{Atoms.}
          The three types of atoms are represented as structures of 2 to 3
          strings, with each string matching an entity identifier in the
          symbol table.

          \subparagraph{Holds Atom.}
            \begin{quote}
              \begin{tabular}[t]{|l|l|l|}
                \hline
                \textbf{Field} & \textbf{Type} & \textbf{Description} \\
                \hline
                $sub$ & string & subject entity identifier \\
                \hline
                $acc$ & string & access right entity identifier \\
                \hline
                $obj$ & string & object entity identifier \\
                \hline
              \end{tabular}
            \end{quote}

          \subparagraph{Member Atom.}
            \begin{quote}
              \begin{tabular}[t]{|l|l|l|}
                \hline
                \textbf{Field} & \textbf{Type} & \textbf{Description} \\
                \hline
                $elt$ & string & single entity identifier \\
                \hline
                $grp$ & string & group entity identifier \\
                \hline
              \end{tabular}
            \end{quote}

          \subparagraph{Subset Atom.}
            \begin{quote}
              \begin{tabular}[t]{|l|l|l|}
                \hline
                \textbf{Field} & \textbf{Type} & \textbf{Description} \\
                \hline
                $grp1$ & string & subgroup entity identifier \\
                \hline
                $grp2$ & string & supergroup entity identifier \\
                \hline
              \end{tabular}
            \end{quote}

        \paragraph{Facts.}
          Facts are represented as a three-element structure composed of the
          following: polymorphic type which can be any of the three atom
          structures above; a type indicator to specify whether the fact is
          holds, member or subset; and a truth flag, to indicate whether the
          atom is classically negated or not ({\em true}, if the fact holds
          and {\em false}, if the fact is classically negated.

          \begin{quote}
            \begin{tabular}[t]{|l|l|l|}
              \hline
              \textbf{Field} & \textbf{Type} & \textbf{Description} \\
              \hline
              $atom$ & atom type & polymorphic structure \\
              \hline
              $type$ & \{h$|$m$|$s\} & holds, member or subset \\ 
              \hline
              $truth$ & boolean & negation indicator \\
              \hline
            \end{tabular}
          \end{quote}
 
        \paragraph{Expressions.}
          Since expressions are simply conjunctions of facts, they are
          represented as an unordered list of fact structures.

        \paragraph{Initial State Facts Table.}
          The initial state facts table is represented as a single list of
          fact structures, or an expression. Each fact in all {\em initially}
          statements are added into the initial state facts table.
 
        \paragraph{Constraint Table.}
          The constraint table is represented as an unordered list of constraint
          structures, with each structure composed of the following:

          \begin{quote}
            \begin{tabular}[t]{|l|l|l|}
              \hline
              \textbf{Field} & \textbf{Type} & \textbf{Description} \\
              \hline
              $exp$ & expression type & consequent \\
              \hline
              $pcond$ & expression type & positive premise \\ 
              \hline
              $ncond$ & expression type & negative premise \\
              \hline
            \end{tabular}
          \end{quote}
 
        \paragraph{Policy Update Table.}
          Another list of structures is the policy update table. Each element
          structure of this table is composed of the following 4 fields:

          \begin{quote}
            \begin{tabular}[t]{|l|l|l|}
              \hline
              \textbf{Field} & \textbf{Type} & \textbf{Description} \\
              \hline
              $name$ & string & update identifier \\
              \hline
              $vlist$ & ordered string list & variables \\
              \hline
              $pre$ & expression type & precondition \\
              \hline
              $post$ & expression type & postcondition \\
              \hline
            \end{tabular}
          \end{quote}
 
        \paragraph{Policy Update Sequence Table.}
 
          The policy update sequence table is an ordered list of sequence
          structures, each with the following elements:
 
          \begin{quote}
            \begin{tabular}[t]{|l|l|l|}
              \hline
              \textbf{Field} & \textbf{Type} & \textbf{Description} \\
              \hline
              $name$ & string & update identifier \\ 
              \hline
              $ilist$ & ordered string list & identifiers \\
              \hline
            \end{tabular}
          \end{quote}

    \subsection{Translation to Normal Logic Program}

      \subsubsection{Grounding of Constraint Variables}

      \subsubsection{Representing Facts in Propositional Form}

        The next step in the translation of language ${\cal L}$ to a normal
        logic program is to represent each fact in propositional form.

        Given the symbol table $st$, we first enumerate all facts defined
        in the policy base:

        \paragraph{Holds}
          \noindent

          \begin{tabular}[t]{|l|l|l|}
            \hline
            $st.s[0]$ & $st.a[0]$ & $st.o[0]$ \\
            \hline
            ... & ... & ... \\
            \hline
            $st.s[0]$ & $st.a[0]$ & $st.o[len(st.o)]$ \\
            \hline
             ... & ... & ... \\
            \hline
            $st.s[0]$ & $st.a[len(st.a)]$ & $st.o[0]$ \\
            \hline
            ... & ... & ... \\
            \hline
            $st.s[0]$ & $st.a[len(st.a)]$ & $st.o[len(st.o)]$ \\
            \hline
            ... & ... & ... \\
            \hline
            $st.s[len(st.s)]$ & $st.a[0]$ & $st.o[0]$ \\
            \hline
            ... & ... & ... \\
            \hline
            $st.s[len(st.s)]$ & $st.a[0]$ & $st.o[len(st.o)]$ \\
            \hline
            ... & ... & ... \\
            \hline
            $st.s[len(st.s)]$ & $st.a[len(st.a)]$ & $st.o[0]$ \\
            \hline
            ... & ... & ... \\
            \hline
            $st.s[len(st.s)]$ & $st.a[len(st.a)]$ & $st.o[len(st.o)]$ \\
            \hline
          \end{tabular}

        \paragraph{Member}
          \noindent

          \begin{tabular}[t]{|l|l|}
            \hline
            % subject
            $st.ss[0]$ & $st.sg[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.ss[0]$ & $st.sg[len(st.sg)]$ \\
            \hline
            ... & ... \\
            \hline
            $st.ss[len(st.ss)]$ & $st.sg[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.ss[len(st.ss)]$ & $st.sg[len(st.sg)]$ \\
            \hline
            % access right
            $st.as[0]$ & $st.ag[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.as[0]$ & $st.ag[len(st.ag)]$ \\
            \hline
            ... & ... \\
            \hline
            $st.as[len(st.as)]$ & $st.ag[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.as[len(st.as)]$ & $st.ag[len(st.ag)]$ \\
            \hline
            % object
            $st.os[0]$ & $st.og[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.os[0]$ & $st.og[len(st.og)]$ \\
            \hline
            ... & ... \\
            \hline
            $st.os[len(st.os)]$ & $st.og[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.os[len(st.os)]$ & $st.og[len(st.og)]$ \\
            \hline
          \end{tabular}

        \paragraph{Subset}
          \noindent

          \begin{tabular}[t]{|l|l|}
            \hline
            % subject
            $st.sg[0]$ & $st.sg[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.sg[0]$ & $st.sg[len(st.sg)]$ \\
            \hline
            ... & ... \\
            \hline
            $st.sg[len(st.sg)]$ & $st.sg[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.sg[len(st.sg)]$ & $st.sg[len(st.sg)]$ \\
            \hline
            % access right
            $st.ag[0]$ & $st.ag[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.ag[0]$ & $st.ag[len(st.ag)]$ \\
            \hline
            ... & ... \\
            \hline
            $st.ag[len(st.ag)]$ & $st.ag[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.ag[len(st.ag)]$ & $st.ag[len(st.ag)]$ \\
            \hline
            % object
            $st.og[0]$ & $st.og[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.og[0]$ & $st.og[len(st.og)]$ \\
            \hline
            ... & ... \\
            \hline
            $st.og[len(st.og)]$ & $st.og[0]$ \\
            \hline
            ... & ... \\
            \hline
            $st.og[len(st.og)]$ & $st.og[len(st.og)]$ \\
            \hline
          \end{tabular}


  \section{Case Study: Web Server Application}

  \section{Conclusion}

    [summary]

    [effectiveness of system/application]

    [future work]

\end{document}
